Cognitive Context: Impact on the Perception of a Robot

### 2015-10-29 main.tex
Removed emphasises with bold face

Robotics researchers often tend to consider that anthropomorphism only describes
a static set of human-like features of a robot (like its shape, its speech
capabilities, facial expressions, etc.).
Following~\citet{fink_anthropomorphism_2012}, we refer to these characteristics
as the \emph{anthropomorphic design} of the robot, and we call
\emph{anthropomorphism} the general \emph{social phenomenon} that emerges from
the (real or imagined) interaction between a non-human agent (a robot in our
case) and a human~\citep{persson_anthropomorphism_2000}.  According
to~\citet{epley_when_2008}, anthropomorphism includes the human perception of
emotional states, motivations, intentions in the non-human agent, and tends to
ascribe those qualities to it. As such, the dynamic and socio-cognitive
dimensions of anthropomorphism are essential to understand the complex bonds
that human users build with artificial agents such as
robots~\citep{lemaignan2014dynamics}. Therefore anthropomorphism appears to be a
useful \emph{proxy} to study these intricate psychological
bonds. Going in that direction, this article introduces \textbf{new ways to assess
anthropomorphic attributions} and provides in particular \textbf{new insights on
the impact of the cognitive context of the interaction on anthropomorphic
attributions}.

@@

After the pre-questionnaire, we invited the participants to an initial
free interaction with a \textbf{powered-off} Nao robot (we did not constraint
the time, and participants played between 2 and 3 minutes with the robot).
The purpose of this pre-interaction was to get the participants acquainted with the
robot shape and kind of motions it could possibly perform, and therefore
mitigate the (visual) novelty effect (so that they do not get surprised seeing
a robot for the first time in the videos).

@@

Our first contribution extends our understanding of the attribution of
human-like characteristics to robot by exploring the role of the interaction
context. As we show, it appears that \textbf{even relatively subtle context
priming may lead to significantly different anthropomorphic perceptions of
the exact same robot, performing the exact same task}, as confirmed both by
distinct gaze patterns, and different reported perceptions in questionnaires.
Interestingly, this effects can already be evidenced after \textbf{very short,
one-way, interactions} (less than 2 minutes).

This leads to a second-order effect that we also evidence here: depending on the
cognitive context, the tendency to attribute human-like characteristics to robot
\textbf{evolves} in significantly different ways. \textbf{After observing a robot in a
context that pre-suppose deeper cognitive capabilities, people increase
their general tendency to anthropomorphize robots} compared to a shallow
cognitive context, even though the robot appearance and visual behaviour is the
exact same.

The second contribution is a methodological one: we present in this article
\textbf{a novel technique to assess anthropomorphic projections that relies on a
biometric measurement (eye-tracking) to compare gaze fixation durations on
the face of the robot}. We cross-validate this new metric with existing,
established, questionnaires. Compared to current techniques (post-hoc
annotations of videos and questionnaires), this new approach is objective,
is less impacted by experimental biases (like the \emph{observer effect},
where the subject's behaviour is impacted by the fact he/she knows that
he/she is observed) and take place during the interaction itself
(\emph{in-the-moment} measurement).

>>>>

Robotics researchers often tend to consider that anthropomorphism only describes
a static set of human-like features of a robot (like its shape, its speech
capabilities, facial expressions, etc.).
Following~\citet{fink_anthropomorphism_2012}, we refer to these characteristics
as the \emph{anthropomorphic design} of the robot, and we call
\emph{anthropomorphism} the general \emph{social phenomenon} that emerges from
the (real or imagined) interaction between a non-human agent (a robot in our
case) and a human~\citep{persson_anthropomorphism_2000}.  According
to~\citet{epley_when_2008}, anthropomorphism includes the human perception of
emotional states, motivations, intentions in the non-human agent, and tends to
ascribe those qualities to it. As such, the dynamic and socio-cognitive
dimensions of anthropomorphism are essential to understand the complex bonds
that human users build with artificial agents such as
robots~\citep{lemaignan2014dynamics}. Therefore anthropomorphism appears to be a
useful \emph{proxy} to study these intricate psychological
bonds. Going in that direction, this article introduces new ways to assess
anthropomorphic attributions and provides in particular new insights on
the impact of the cognitive context of the interaction on anthropomorphic
attributions.

@@

After the pre-questionnaire, we invited the participants to an initial
free interaction with a powered-off Nao robot (we did not constraint
the time, and participants played between 2 and 3 minutes with the robot).
The purpose of this pre-interaction was to get the participants acquainted with the
robot shape and kind of motions it could possibly perform, and therefore
mitigate the (visual) novelty effect (so that they do not get surprised seeing
a robot for the first time in the videos).

@@

Our first contribution extends our understanding of the attribution of
human-like characteristics to robot by exploring the role of the interaction
context. As we show, it appears that even relatively subtle context
priming may lead to significantly different anthropomorphic perceptions of
the exact same robot, performing the exact same task, as confirmed both by
distinct gaze patterns, and different reported perceptions in questionnaires.
Interestingly, this effects can already be evidenced after very short,
one-way, interactions (less than 2 minutes).

This leads to a second-order effect that we also evidence here: depending on the
cognitive context, the tendency to attribute human-like characteristics to robot
evolves in significantly different ways. After observing a robot in a
context that pre-suppose deeper cognitive capabilities, people increase
their general tendency to anthropomorphize robots compared to a shallow
cognitive context, even though the robot appearance and visual behaviour is the
exact same.

The second contribution is a methodological one: we present in this article
a novel technique to assess anthropomorphic projections that relies on a
biometric measurement (eye-tracking) to compare gaze fixation durations on
the face of the robot. We cross-validate this new metric with existing,
established, questionnaires. Compared to current techniques (post-hoc
annotations of videos and questionnaires), this new approach is objective,
is less impacted by experimental biases (like the \emph{observer effect},
where the subject's behaviour is impacted by the fact he/she knows that
he/she is observed) and take place during the interaction itself
(\emph{in-the-moment} measurement).
